**OptiFit: a fast method for fitting amplicon sequences to existing OTUs**

### 2021-01-22 paper/manuscript.Rmd
outline edits

# TITLE GOES HERE

@@

**Running title:** Optimal resolution

@@

Courtney R. Armour, (ande? begum? etc?), Patrick D. Schloss ${^\dagger}$

@@

- CRC is one of the most common cancers and a leading cause of cancer related death
- There is evidence that the microbiome has a role in CRC development/progression and could be useful for biomarker detection and diagnostics.
- Begum et al (mBio 2020) recently demonstrated effective application of machine learning (ML) to microbiome based classification problems and developed a framework for applying ML practices in a more reproducible way (mikropml).
- A common question when applying ML methods to microbiome data is which method and taxonomic level should be use.
- This analysis utilizes the reproducible framework developed by Begum et al to quantify which ML method and taxonomic level produce the best performing classifier for CRC data.

@@

## Results
- Across the five ML methods tested, model performance tends to increase with taxnomic level usually peaking around genus level and dropping off slightly with ASVs.
- Random forest was consistently the top performer at most taxonomic levels.
  - RF might be more appropriate for microbiome analysis since its more suitable for zero inflated data (need to look into literature)
- Within the RF model, the highest AUCs were observed for family, genus, and otu level data with no significant difference between the three (Figure 1). ASV performed significantly lower than OTU but not family and genus levels.

@@

## Materials and Methods
- data from prior study {baxter}
- mikropml package
- pvalues as previously described {begum}

>>>>

# Optimal Resolution for Microbiome-Based Classification of Colorectal Cancer

@@

**Running title:** Optimal Resolution

@@

Courtney R. Armour, (ande? begum?), Patrick D. Schloss ${^\dagger}$

@@

- CRC is one of the most common cancers and a leading cause of cancer related death
- There is evidence that the microbiome has a role in CRC development/progression and could be useful for biomarker detection and diagnostics.
- Begum et al (mBio 2020) recently demonstrated effective application of machine learning (ML) to microbiome based classification problems and developed a framework for applying ML practices in a more reproducible way (mikropml).
- A common question when applying ML methods to microbiome data is which method and taxonomic level is optimal.
- This analysis utilizes the reproducible framework developed by Begum et al to quantify which ML method and taxonomic level produce the best performing classifier for CRC data.

@@

## Results
- Across the five ML methods tested, model performance tends to increase with taxonomic level usually peaking around genus/otu level and dropping off slightly with ASVs.
- Random forest was consistently the top performer at most taxonomic levels.
  - RF might be more appropriate for microbiome analysis since its more suitable for zero inflated data (need to look into literature)
- Within the RF model, the highest AUCs were observed for family, genus, and otu level data with no significant difference between the three (Figure 1). ASV performed significantly lower than OTU but not family and genus levels.

@@

## Materials and Methods
- data from prior study {baxter}
- mikropml package
  - preprocessing
  - hp tuning
  - n seeds
- pvalues as previously described {begum}

### 2021-01-22 paper/manuscript.tex
outline edits

\hypertarget{title-goes-here}{%
\section{TITLE GOES HERE}\label{title-goes-here}}

@@

\textbf{Running title:} Optimal resolution

@@

Courtney R. Armour, (ande? begum? etc?), Patrick D. Schloss
\({^\dagger}\)

@@

\begin{itemize}
\tightlist
\item
  CRC is one of the most common cancers and a leading cause of cancer
  related death
\item
  There is evidence that the microbiome has a role in CRC
  development/progression and could be useful for biomarker detection
  and diagnostics.
\item
  Begum et al (mBio 2020) recently demonstrated effective application of
  machine learning (ML) to microbiome based classification problems and
  developed a framework for applying ML practices in a more reproducible
  way (mikropml).
\item
  A common question when applying ML methods to microbiome data is which
  method and taxonomic level should be use.
\item
  This analysis utilizes the reproducible framework developed by Begum
  et al to quantify which ML method and taxonomic level produce the best
  performing classifier for CRC data.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Across the five ML methods tested, model performance tends to increase
  with taxnomic level usually peaking around genus level and dropping
  off slightly with ASVs.
\item
  Random forest was consistently the top performer at most taxonomic
  levels.

>>>>

\hypertarget{optimal-resolution-for-microbiome-based-classification-of-colorectal-cancer}{%
\section{Optimal Resolution for Microbiome-Based Classification of
Colorectal
Cancer}\label{optimal-resolution-for-microbiome-based-classification-of-colorectal-cancer}}

@@

\textbf{Running title:} Optimal Resolution

@@

Courtney R. Armour, (ande? begum?), Patrick D. Schloss \({^\dagger}\)

@@

\begin{itemize}
\tightlist
\item
  CRC is one of the most common cancers and a leading cause of cancer
  related death
\item
  There is evidence that the microbiome has a role in CRC
  development/progression and could be useful for biomarker detection
  and diagnostics.
\item
  Begum et al (mBio 2020) recently demonstrated effective application of
  machine learning (ML) to microbiome based classification problems and
  developed a framework for applying ML practices in a more reproducible
  way (mikropml).
\item
  A common question when applying ML methods to microbiome data is which
  method and taxonomic level is optimal.
\item
  This analysis utilizes the reproducible framework developed by Begum
  et al to quantify which ML method and taxonomic level produce the best
  performing classifier for CRC data.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Across the five ML methods tested, model performance tends to increase
  with taxonomic level usually peaking around genus/otu level and
  dropping off slightly with ASVs.
\item
  Random forest was consistently the top performer at most taxonomic
  levels.
